\documentclass{article}%ctex
\input{~/code/math_commands.tex}




\title{\huge R3\\
\normalsize}
\author{Xuanxi Zhang}
\begin{document}
\maketitle
\section{}
Consider the following letter-envelope matching problem. There are $n$ letters labeled 1 through $n$ and also $n$ envelopes labeled 1 through $n$. The letters are permuted randomly into the envelopes (possibly by a drunk person), one letter per envelope, so that all permutations are equally likely. Let $X_n$ be the number of letters that are placed into their matching envelopes.
\begin{enumerate}
    \item Find the probability $\mathbb{P}\left(X_n=0\right)$. (Hint: inclusion-exclusion principle)
    \item Find the limiting distribution of $X_n$ as $n \rightarrow \infty$. Is it a normal distribution?

\end{enumerate}
For reference, here is the inclusion-exclusion principle we proved last class:
$$
\mathbb{P}\left(\bigcup_{i=1}^n A_i\right)=\sum_{k=1}^n(-1)^{k-1} \sum_{I \subseteq\{1,2, \cdots, n\}:|I|=k} \mathbb{P}\left(\bigcap_{i \in I} A_i\right)
$$


\section{}
$\{X_i\}_{i=1}^n$ are Bernoulli random variables with parameter $1/2$ and they are not necessarily independent.
Show that
$$
\E\left(\frac{1}{n}\sum_{i=1}^n X_i>\frac{1}{4}\right)\geq \frac{1}{3}
$$


\section{7.3} Which of the following is a random variable?
\begin{enumerate}
    \item population mean
    \item population size, $N$
    \item sample size, $n$
    \item sample mean
    \item variance of sample mean
    \item largest value in sample
    \item population variance
    \item estimated variance of sample mean
\end{enumerate}

\section{7.10}
 True or False? If a sample from a population is large, a histogram of the values in the sample will be approximately normal, even if the population is not normal.



\section{7.9} In a random sample of 1,500 voters, $55 \%$ said they planned to vote for a certain proposition, and $45 \%$ said they would vote against it. The estimated margin of victory for the proposition is thus $55 \%-45 \%=10 \%$. What is the standard error of this estimated margin? What is a $95 \%$ confidence interval for the margin? (You'll need a calculator or a device that can serve as one.)

\section{True or False}

\begin{enumerate}
    \item Which is more likely, 9 heads in 10 tosses of a fair coin, or 18 heads in 20 tosses? (You don't need to calculate the exact probabilities.)
    \item In order to halve the width of a $95 \%$ confidence interval for a mean, by what factor should the sample size be increased?
    \item True or False? A 95\% confidence interval contains the sample mean with probability 0.95 .
    \item Suppose you survey 100 people and ask them to rate their opinion of a product from -5 to 5 , then average their ratings and divide by the standard error. If the population mean of the ratings is 0 , approximately what type of distribution does the square of this scaled average follow?
    \item Suppose we measure the heights of 10 NYU students. If we wanted to find a confidence interval for the average height of an NYU student, we'd look in a table for the CDF of what distribution?
\end{enumerate}


%\bibliographystyle{plainnat}
%\bibliography{ref}
\end{document}